---
title: "All Countries Environmental Data"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---


```{r include=FALSE, results='hide'}
# Installing required packages
# sudo apt-get install libcairo2-dev
# sudo apt-get install libxml2-dev

devtools::install_github('HerveAbdi/PTCA4CATA')
library(corrplot)
library(ExPosition)
library(InPosition)
library(animation)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(gplots)
library(PTCA4CATA)
library(reshape2)

# Source the function file:
#
# install.packages('TExPosition')
# install.packages('MExPosition')
library(MExPosition)

```


## Method: MFA


## Dataset

```{r include=FALSE, results='hide'}
country_env_df = read.csv('../dataset/World_countries_env_vars.csv')
country_env_df_for_corr = country_env_df[, !names(country_env_df) %in% c('Country')]

country_env_df$Happiness_Rank = cut(country_env_df$Happiness_Rank,breaks = c(0, 39,118, 159), right=TRUE, labels = c('VH' , 'H', 'U'),ordered_result = TRUE)
```

* Data: Measurements of environment conditions in Countries
* Rows: There are 137 observations, 1 for each country.
* Columns: Columns: Total 2 Tables – All Rain and All Temp

* Structure of Data

```{r data_summary}
str(country_env_df)
```


* Research Question

Which variables in Rain and Temperature contribute most towards happiness

## Analysis 

```{r echo=FALSE}

country_env_df_for_mfa = country_env_df[, !names(country_env_df) %in% c('Country', 'Happiness_Rank')]

rownames(country_env_df_for_corr) <- paste0(country_env_df[,'Country'], '(' ,country_env_df[,'Happiness_Rank'], ')')
rownames(country_env_df_for_mfa) <- paste0(country_env_df[,'Country'], '(' ,country_env_df[,'Happiness_Rank'], ')')

```


### Correlation Plot
Visually analyze multicollinearity between all varaibles in Rain and Temperature tables.  

```{r}
column.design = rep('Others', length.out = ncol(country_env_df_for_mfa) )

column.design[ grep('rain',colnames(country_env_df_for_mfa)) ] = 'Rain'
column.design[ grep('temp',colnames(country_env_df_for_mfa)) ] = 'Temp'
column.design = t(as.matrix(column.design))
colnames(column.design) = colnames(country_env_df_for_mfa)
rownames(column.design) = 'Type'
```

### Identify Latent Components

#### MFA 

```{r pca, echo=FALSE, results='hide'}

resMFA <- mpMFA(country_env_df_for_mfa,  as.matrix(column.design), DESIGN = country_env_df$Happiness_Rank,graphs = TRUE, make.columndesign.nominal = TRUE, make.design.nominal = TRUE)

#resPerm4PLSC <- mpM (rain, # First Data matrix 
#                          temp, # Second Data matrix
#                          nIter = 1000 # How mny iterations
#                          )
# to see what results we have


```

Now we have Latent Variables and Saliences. 
* Latent Variables are the new Data points w.r.t. correlation between both the tables. Latent Variables exists for each table.
* Saliences represent correlation between variables of each table. 

#### Scree Plot
Gives amount of information explained by corresponding component. Gives an intuition to decide which components best represent data in order to answer the research question.

P.S. The most contribution component may not always be most useful for a given research question. 


```{r}


PTCA4CATA::PlotScree(ev = resMFA$mexPosition.Data$Table$eigs,
                      title = 'SCREE Plot',
                      #p.ev = resPerm4PLSC$pEigenvalues,
                      plotKaiser = TRUE
)

```

#### Latent Variables 

Lets visualize happiness categories for Components 1 of each table

##### Factor Scores
```{r echo=FALSE}

comp1 = resMFA$mexPosition.Data$Table$fi

list_color <- c('U' = 'burlywood3', 'H' = 'royalblue1', 'VH' = 'springgreen4')

lx_col <- recode (country_env_df$Happiness_Rank, `U` = 'burlywood3', `H` = 'royalblue1', `VH` = 'springgreen4')

country_env_mean_df = aggregate(comp1,by = list(country_env_df$Happiness_Rank), mean)
rownames(country_env_mean_df) <- c('Very Happy', 'Happy', 'Unhappy')

country_env_mean_df <- country_env_mean_df[,2:3]

country_factor_map <- PTCA4CATA::createFactorMap(comp1, title='', 
                                                 col.points = lx_col,
                                                 #col.labels = country_env_df$Happiness_Rank,
                                                 
                                                 display.labels = FALSE)

country_factor_mean_map <- PTCA4CATA::createFactorMap(country_env_mean_df, title='', 
                                                 col.points = c('springgreen4', 'royalblue1', 'burlywood3'),
                                                 col.labels = c('springgreen4', 'royalblue1', 'burlywood3'),
                                                 cex = 8,alpha.points = 0.8,
                                                 display.labels = TRUE)

country_label4Map <- PTCA4CATA::createxyLabels.gen(1,2,lambda = resMFA$mexPosition.Data$Table$eigs,  tau = resMFA$mexPosition.Data$Table$t ,axisName = "Table") 


country_map = country_factor_map$zeMap + country_label4Map + country_factor_mean_map$zeMap_dots + country_factor_mean_map$zeMap_text + ggtitle('Component 1')

print(country_map)

```

#### Loadings

##### Per Component Plot
```{r echo=FALSE}

PrettyBarPlot2(resMFA$mexPosition.Data$Table$Q[,1],
              threshold = 0,
              font.size = 5,
              main = 'Salience for Rain',
              horizontal = TRUE,
              angle.text = 90,
              ylim=c(-3,3)
              )

```

```{r echo=FALSE}

PrettyBarPlot2(resMFA$mexPosition.Data$Table$Q[,2],
              threshold = 0,
              font.size = 5,
              main = 'Salience for Rain',
              horizontal = TRUE,
              angle.text = 90,
              ylim=c(-3,3)
              )

```

#### Correlation Circle

```{r}

loadings_2 = cor(country_env_df_for_mfa, resMFA$mexPosition.Data$Table$fi)

col_palate = brewer.pal(n = 12, name='Set3')

col4J = vector('list', nrow(loadings_2))
col4J[grep('rain',rownames(loadings_2))] = col_palate[1]
col4J[grep('temp',rownames(loadings_2))] = 'red' #col_palate[2]
col4J[sapply(col4J, 'is.null')] = rep(col_palate[4], length.out = sum(sapply(col4J, 'is.null')))


loadings_map <- PTCA4CATA::createFactorMap(loadings_2, 
                          col.points = col4J, 
                          col.labels = col4J, 
                          axis1=1,
                          axis2=2,
                          constraints = list(minx = -1, miny = -1, maxx = 1 , maxy = 1)) 
  
  country_label4Map <- PTCA4CATA::createxyLabels.gen(1,2,lambda = resMFA$mexPosition.Data$Table$eigs, tau = resMFA$mexPosition.Data$Table$t) 
  
  corr_map <- loadings_map$zeMap_background  +  country_label4Map + PTCA4CATA::addCircleOfCor() +
              loadings_map$zeMap_text + PTCA4CATA::addArrows(loadings_2, color = col4J) 
  
  print(corr_map)

```

### Conclusion

Here Component 2 seems to best seperate Happiness levels. Let's compare Component 2 for both tables.

* Table 1 & 2 Component 2
  + Latent Variables: Very Happy vs Unhappy (for Rain and Temperature both)
  + Salience:
    * Rain: It seems dryness and wetness at a montly scale have more effect than coldness or yearly patterns.
    * Temperature: All temperature variations at a monthly and yearly scale seems to impact happiness.
  

