---
title: "All Countries Environmental Data"
output:
  word_document: default
  html_notebook: default
  pdf_document: default
---


```{r include=FALSE, results='hide'}
# Installing required packages
# sudo apt-get install libcairo2-dev
# sudo apt-get install libxml2-dev

devtools::install_github('HerveAbdi/PTCA4CATA')
library(corrplot)
library(ExPosition)
library(InPosition)
library(animation)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(gplots)
library(PTCA4CATA)
library(flextable)
library(TInPosition)
library(rvg)
library(dummies)
library(TExPosition)
library(TInPosition)
rm(list = ls())
```

# Barycentric Discriminant Analysis

## Method: BADA

<Will update after class>


```{r include=FALSE, results='hide'}
country_env_df = read.csv('../dataset/World_countries_env_vars.csv')
country_env_df_for_corr = country_env_df[, !names(country_env_df) %in% c('Country')]

country_env_df$Happiness_Rank = cut(country_env_df$Happiness_Rank,breaks = c(0, 39,118, 159), right=TRUE, labels = c('VH' , 'H', 'U'),ordered_result = TRUE)

str(country_env_df)
```

```{r echo=FALSE}

country_env_df_for_pca = country_env_df[, !names(country_env_df) %in% c('Country', 'Happiness_Rank')]

rownames(country_env_df_for_corr) <- paste0(country_env_df[,'Country'], '(' ,country_env_df[,'Happiness_Rank'], ')')
rownames(country_env_df_for_pca) <- paste0(country_env_df[,'Country'], '(' ,country_env_df[,'Happiness_Rank'], ')')

```


## Heatmap
Visually analyze multicollinearity in the system.  

```{r include=FALSE}
corr_result = cor(country_env_df_for_corr)
corrplot(corr_result,order = 'hclust', addrect = 7)
```

```{r}

happiness_dummies = as.data.frame(dummy(country_env_df$Happiness_Rank))
colnames(happiness_dummies) <- c('Happiest', 'Happier', 'Happy' )

#heatmap(t(happiness_dummies) %*% as.matrix(country_env_df_for_pca))
```

```{r}
heatmap.2(t(happiness_dummies) %*% as.matrix(country_env_df_for_pca), col = rev(heat.colors(16)), dendrogram = 'none', trace = 'none', margins = c(8, 16)) #tracecol=NA)
```


```{r bada, include=FALSE}

country_env_pca <- tepBADA(DATA = country_env_df_for_pca, center = TRUE, scale = 'SS1', DESIGN = country_env_df$Happiness_Rank, graphs = FALSE)
country_env_pca_inf <- tepBADA.inference.battery(DATA = country_env_df_for_pca, center = TRUE, scale = 'SS1', DESIGN = country_env_df$Happiness_Rank, graphs = FALSE)
```

## Scree Plot
Gives amount of information explained by corresponding component. Gives an intuition to decide which components best represent data in order to answer the research question.

P.S. The most contribution component may not always be most useful for a given research question. 


```{r echo=FALSE}

PTCA4CATA::PlotScree(ev = country_env_pca$TExPosition.Data$eigs,
                      #p.ev =  country_env_pca_inf$Inference.Data$components$p.vals,
                      title = 'SCREE Plot',
                      plotKaiser = TRUE
)

```

## Factor Scores 


```{r echo=FALSE}


list_color <- c('#305ABF' = 'burlywood3', '#84BF30' = 'royalblue1', '#BF30AD' = 'springgreen4')
happy_levels <- c('Happy', 'Happier', 'Happiest')

country_env_pca$Plotting.Data$fii.col[,1] <- recode (country_env_pca$Plotting.Data$fii.col, !!!list_color)

groups_of_data <- vector('list', 3)

col = country_env_pca$Plotting.Data$fii.col
fi = country_env_pca$TExPosition.Data$fii

j = 1
color_filter = col == unique(col)[1]
groups_of_data[[j]] = color_filter
j = j + 1

color_filter = col == unique(col)[2]
groups_of_data[[j]] = color_filter
j = j + 1

color_filter = col == unique(col)[3]
groups_of_data[[j]] = color_filter
j = j + 1

#country_env_pca_mean = aggregate(x=country_env_pca$TExPosition.Data$fi, by = split(col, col(col)), FUN = mean)
rownames(country_env_pca$TExPosition.Data$fi) <- happy_levels

#country_env_pca_mean <- country_env_pca_mean[-1]

```


```{r echo=FALSE}

axis1 = 1
axis2 = 2
country_factor_map <- PTCA4CATA::createFactorMap(country_env_pca$TExPosition.Data$fii, title='', 
                                                 col.points = country_env_pca$Plotting.Data$fii.col,
                                                 col.labels = country_env_pca$Plotting.Data$fii.col,
                                                 axis1 = axis1,
                                                 axis2 = axis2,
                                                 display.labels = FALSE)

country_factor_map_mean <- PTCA4CATA::createFactorMap(country_env_pca$TExPosition.Data$fi,
                                                 col.points = c('burlywood3', 'royalblue1', 'springgreen4'),
                                                 col.labels = c('burlywood3', 'royalblue1', 'springgreen4'),
                                                 axis1 = axis1,
                                                 axis2 = axis2,
                                                 display.labels = TRUE,
                                                 cex = 8,alpha.points = 0.8)

country_label4Map <- PTCA4CATA::createxyLabels.gen(axis1,axis2,lambda = country_env_pca$TExPosition.Data$eigs, tau = country_env_pca$TExPosition.Data$t) 



country_map = country_factor_map$zeMap + country_factor_map_mean$zeMap_dots + country_factor_map_mean$zeMap_text +country_label4Map


BootCube <- country_env_pca_inf$Inference.Data$boot.data$fi.boot.data$boots
dimnames(BootCube)[[2]] <- c("Dimension 1","Dimension 2")

# use function MakeCIEllipses from package PTCA4CATA
GraphElli <- MakeCIEllipses(BootCube[,1:2,],
                  names.of.factors = c("Dimension 1","Dimension 2"),
                  col = unique(country_env_pca$Plotting.Data$fii.col),
                  p.level = .95
)
#country_map = country_map + GraphElli

print(country_map + GraphElli)


```


* With Tolerance Interval

```{r echo=FALSE}

axis1 = 1
axis2 = 2

GraphTI.Hull <- MakeToleranceIntervals(country_env_pca$TExPosition.Data$fii[,c(axis1, axis2)],
                                                  design = country_env_df$Happiness_Rank,
                                                  # line below is needed
                                                  names.of.factors =  c("Dim1","Dim2"), # needed 
                                                  col = unique(country_env_pca$Plotting.Data$fii.col),
                                                  line.size = .50, 
                                                  line.type = 3,
                                                  alpha.ellipse = .2,
                                                  alpha.line    = .4,
                                                  p.level       = .75)

print(country_map + GraphTI.Hull)

```


## Loadings

```{r echo=FALSE}


axis1 = 1
axis2 = 2

col_palate = brewer.pal(n = 12, name='Set3')

col4J = vector('list', nrow(country_env_pca$TExPosition.Data$fj))
col4J[grep('rain',rownames(country_env_pca$TExPosition.Data$fj))] = col_palate[1]
col4J[grep('temp',rownames(country_env_pca$TExPosition.Data$fj))] = 'red' #col_palate[2]
col4J[sapply(col4J, 'is.null')] = col_palate[3:11]

loadings_2 <- cor(country_env_df_for_pca, country_env_pca$TExPosition.Data$fii)

loadings_map <- PTCA4CATA::createFactorMap(loadings_2, 
                      col.points = col4J, 
                      col.labels = col4J, 
                      axis1=axis1,
                      axis2=axis2,
                      constraints = list(minx = -1, miny = -1, maxx = 1 , maxy = 1)) 

country_label4Map <- PTCA4CATA::createxyLabels.gen(axis1,axis2,lambda = country_env_pca$TExPosition.Data$eigs, tau = country_env_pca$TExPosition.Data$t) 

corr_map <- loadings_map$zeMap_background + PTCA4CATA::addCircleOfCor() +
          loadings_map$zeMap_text +
          PTCA4CATA::addArrows(loadings_2, color = col4J) + country_label4Map

print(corr_map)

```


## Most Contributing Variables

* With Bootstrap Ratio

```{r echo=FALSE}

BR <- country_env_pca_inf$Inference.Data$boot.data$fj.boot.data$tests$boot.ratios

for (i in c(1, 2)) {
  laDim = i
  ba001.BR1 <- PrettyBarPlot2(BR[,laDim],
                              threshold = 2,
                              font.size = 5,
                              #color4bar = gplots::col2hex(col4J), # we need hex code
                              main = paste0('Bootstrap ratio ',laDim),
                              ylab = 'Bootstrap ratios'
                              #ylim = c(1.2*min(BR[,laDim]), 1.2*max(BR[,laDim]))
  )
  print(ba001.BR1)
}
```


## Permutation Test

```{r echo=FALSE}
for (i in c(1, 2)) {
zeDim = i
pH1 <- prettyHist(
  distribution = country_env_pca_inf$Inference.Data$components$eigs.perm[,zeDim], 
  observed = country_env_pca_inf$Fixed.Data$TExPosition.Data$eigs[zeDim], 
  xlim = c(0, country_env_pca_inf$Fixed.Data$TExPosition.Data$eigs[zeDim]+2), # needs to be set by hand
  breaks = 20,
  border = "black", 
  main = paste0("BADA: Permutation Test for Eigenvalue ",zeDim),
  xlab = paste0("Eigenvalue ",zeDim), 
  ylab = "", 
  counts = FALSE, 
  cutoffs = c(0.025, 0.975))
}
```

## Parallet Test

```{r echo=FALSE}
country_env_pca_mc <- data4PCCAR::monteCarlo.eigen(X = country_env_df_for_pca, nIter = 1000)
for (i in c(1, 2)) {
  zeDim = i
  pH1.p <- prettyHist(country_env_pca_mc$rand.eigs[,zeDim], 
                    observed = country_env_pca_mc$fixed.eigs[zeDim], 
                    xlim = c(0, country_env_pca_mc$fixed.eigs[zeDim]+2), # needs to set by hand
                    breaks = 20,
                    border = "black", 
                    main = paste0("Monte Carlo (Parallel) Test for Eigenvalue ",zeDim),
                    xlab = paste0("Eigenvalue ",zeDim), 
                    ylab = "", 
                    counts = FALSE, 
                    cutoffs = c(0.025,0.975))
}

```

## Bootstrap Test

```{r echo=FALSE}

#country_env_pca_br <- PTCA4CATA::Boot4Mean(country_env_pca$ExPosition.Data$fi, design = country_env_df$Happiness_Rank, niter=100, suppressProgressBar = FALSE)
country_env_pca_bs <- data4PCCAR::boot.eigen(X = country_env_df_for_pca, nIter = 1000)

for (i in c(1, 2)) {
  zeDim = i
  prettyHist(country_env_pca_bs$boot.eigs[,zeDim], 
                    observed = country_env_pca_bs$fixed.eigs[zeDim], 
                    xlim = c(0, country_env_pca_bs$fixed.eigs[zeDim]+2), # needs to set by hand
                    breaks = 20,
                    border = "black", 
                    main = paste0("Bootstrapped distribution for Eigenvalue ",zeDim),
                    xlab = paste0("Eigenvalue ",zeDim), 
                    ylab = "", 
                    counts = FALSE, 
                    cutoffs = c(0.025, 0.975))
}
```

## Conclusion

* Component 1: 
  + Rows: Normal & Happy
  + Columns: Cloudiness & Rain vs Cropland, Aspect, Elevation
  + Interpret: People in countries with more Cloudiness, Trees and Rain tends to be happier.

* Component 7: 
  + Rows: Happy & Unhappy
  + Columns: Temp and Rain vs Accessibility and Cropland
  + Interpret: Rain and Temp seems to be main reason for unhappiness and Cropland is important for Happiness.

* Component 9: 
  + Rows: Happy & Very Happy
  + Columns: Temp vs Rain
  + Interpret: Rain and Temp seems to be main reason for Happiness. *This contradicts with Component 7 and 1*.



