---
title: "All Countries Environmental Data"
output:
  word_document: default
  html_notebook: default
  pdf_document: default
---


```{r include=FALSE, results='hide'}
# Installing required packages
# sudo apt-get install libcairo2-dev
# sudo apt-get install libxml2-dev

devtools::install_github('HerveAbdi/PTCA4CATA')
library(corrplot)
library(ExPosition)
library(InPosition)
library(animation)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(gplots)
library(PTCA4CATA)
library(purrr)
library(tidyr)
library(TExPosition)
library(TInPosition)
library(dummies)
suppressMessages(library(flextable))
suppressMessages(library(rvg))
# use the last version of PTCA

# also install data4PCCAR last version
devtools::install_github('HerveAbdi/data4PCCAR')
library(data4PCCAR)
library(cowplot)

```


## Method: DiCA



## Dataset

```{r include=FALSE, results='hide'}
country_env_df = read.csv('../dataset/World_countries_env_vars.csv')
country_env_df_for_corr = country_env_df[, !names(country_env_df) %in% c('Country')]

country_env_df$Happiness_Rank = cut(country_env_df$Happiness_Rank,breaks = c(0, 39,118, 159), right=TRUE, labels = c('VH' , 'H', 'U'),ordered_result = TRUE)
```

* Data: Measurements of environment conditions in Countries
* Rows: There are 137 observations, 1 for each country.
* Columns: Total 29 variables
 + Qualitative: Country (nominal), Happiness (Ordinal).
 + Quantitative: Aspect, Slope Crop Land, Tree Canopy Wind Cloud & Multiple variables for Temp & Rain

* Structure of Data

```{r data_summary}
str(country_env_df)
```


* Research Question

How do the 137 countries differ on these variables?

## Analysis 

Let's observe the distribution of each variables to get an intuition of how we can bin these variables. It's important to have nearly equal number of observations in the each bin and to try to cut the variables in a way to so that each new binned distribution is nearly Gaussian. We can also verify that our binning is appropiate by calculating Spearman Correlation for each of original variable and binned variable, the correlation coefficient should be close to 1.

```{r echo=FALSE}

country_env_df_for_pca = country_env_df[, !names(country_env_df) %in% c('Country', 'Happiness_Rank')]

rownames(country_env_df_for_corr) <- paste0(country_env_df[,'Country'], '(' ,country_env_df[,'Happiness_Rank'], ')')
rownames(country_env_df_for_pca) <- paste0(country_env_df[,'Country'], '(' ,country_env_df[,'Happiness_Rank'], ')')

```

### Density plot
```{r echo=FALSE}

cut_points = list()

cut_points$accessibility_to_cities <- c(200,500)
cut_points$aspect <- c(160,180)
cut_points$cloudiness <- c(150,220)
cut_points$cropland_cover <- c(20,40)
cut_points$elevation <- c(200,800)
cut_points$isothermality <- c(40,65)
cut_points$rain_coldestQuart <- c(150, 300)
cut_points$rain_driestMonth <- c(15, 40)
cut_points$rain_driestQuart <- c(70,150)
cut_points$rain_mean_annual <- c(500,1500)
cut_points$rain_seasonailty <- c(50, 80)
cut_points$rain_warmestQuart <- c(100,300)
cut_points$rain_wettestMonth <- c(100, 300)
cut_points$rain_wettestQuart <- c(250, 700)
cut_points$slope <- c(0.7, 1.5)
cut_points$temp_annual_range <- c(20,30)
cut_points$temp_coldestQuart <- c(0,15)
cut_points$temp_diurnal_range <- c(10,14)
cut_points$temp_driestQuart <- c(10,20)
cut_points$temp_max_warmestMonth <- c(25,35)
cut_points$temp_mean_annual <- c(12,25)
cut_points$temp_min_coldestMonth <- c(0,10)
cut_points$temp_seasonality <- c(30, 70)
cut_points$temp_warmestQuart <- c(20, 25)
cut_points$temp_wettestQuart <- c(15, 25)
cut_points$tree_canopy_cover <- c(10, 30)
cut_points$wind <- c(2,3,4)
```


```{r}

plot_bins <- function(df, col) {
  plot_col <- ggplot(df) + aes(df[,col]) +  geom_density() + 
    geom_vline(xintercept = cut_points[[col]], color = 'red') +
    xlab(col) + theme_minimal() + ylab(NULL) +
  theme(axis.text.y = element_blank())
  return(plot_col)
}

bin_hist_list <- lapply(colnames(country_env_df_for_pca), function (col) plot_bins(country_env_df_for_pca, col))

plot_grid(plotlist=bin_hist_list)

```


### Binning

Structure of Data after binning based on above observation.

```{r echo=FALSE}

cut_r <- function(x, b = 4 ) {
  c <- cut(x,breaks = b, include.lowest = TRUE, labels = FALSE)
  return (c) #(factor(c))
}

country_env_df_for_mca = mapply(function(col, col_name) cut_r(col, c(min(col), cut_points[[col_name]], max(col))) , country_env_df_for_pca, colnames(country_env_df_for_pca)  )
country_env_df_for_mca <- as.data.frame(country_env_df_for_mca)
str(country_env_df_for_mca)
```

### Spearman Correlation

Let's observe correlation between original data and binned data to make sure that neither the correlation ceofficient is too low or perfect.
```{r}

cor_spear <- mapply(function(x,y) cor(x, as.integer(y),method = "spearman"), country_env_df_for_pca, country_env_df_for_mca)
#columns = colnames(country_env_df_for_pca)
#cor_df <- data.frame(col = columns, corr = cor_spear)
cor_p <- as.data.frame(cor_spear)

ggplot(data=cor_p, aes(x=rownames(cor_p), y=cor_p$cor_spear)) +
    geom_bar(stat="identity") + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) +
  xlab("") + ylab("Spearmen Correlation") + ylim(0, 1)
  

```



```{r pca, echo=FALSE, results='hide', message=FALSE}

happiness_dummies = as.data.frame(dummy(country_env_df$Happiness_Rank))
colnames(happiness_dummies) <- c('Happiest', 'Happier', 'Happy' )


country_env_mca <- tepDICA(DATA = as.matrix(country_env_df_for_mca), DESIGN = happiness_dummies, graphs = FALSE, make_data_nominal = TRUE)

country_env_mca_inf <- tepDICA.inference.battery(DATA = as.matrix(country_env_df_for_mca), DESIGN = happiness_dummies, graphs = FALSE, test.iters=50, make_design_nominal=FALSE)

```



### Correlation Plot
Visually analyze multicollinearity in the system of the original data

```{r}

heatmap.2(t(happiness_dummies) %*% as.matrix(country_env_df_for_mca), col = rev(heat.colors(16)), dendrogram = 'none', trace = 'none', margins = c(8, 16)) #tracecol=NA)

```

### Heatmap

For binned data.

```{r}

corrMatBurt.list <- phi2Mat4BurtTable(country_env_df_for_mca)
corr4MCA <- corrplot.mixed(as.matrix(corrMatBurt.list$phi2.mat,
                                     title = "Phi2: (squared) Correlation Map for MCA"))
a0001a.corMat.phi2 <- recordPlot()

```



## Identify Latent Components - MCA (With Inference)


### Scree Plot
Gives amount of information explained by corresponding component. Gives an intuition to decide which components best represent data in order to answer the research question.

P.S. The most contribution component may not always be most useful for a given research question. 


```{r}

PTCA4CATA::PlotScree(ev = country_env_mca$TExPosition.Data$eigs,
                      p.ev =  country_env_mca_inf$Inference.Data$components$p.vals,
                      title = 'SCREE Plot',
                      plotKaiser = TRUE
)

```

### Factor Scores 

**With Confidence Interval**
```{r echo=FALSE}
BootCube.Gr <- PTCA4CATA::Boot4Mean(country_env_mca$TExPosition.Data$fii, 
                                    design = country_env_df$Happiness_Rank,
                                    niter = 1000,
                                    suppressProgressBar = TRUE)

  
axis1 = 1
axis2 = 2

  country_factor_map <- PTCA4CATA::createFactorMap(country_env_mca$TExPosition.Data$fii, title='', 
                                                 col.points = country_env_mca$Plotting.Data$fii.col,
                                                 col.labels = country_env_mca$Plotting.Data$fii.col,
                                                 axis1 = axis1,
                                                 axis2 = axis2,
                                                 display.labels = FALSE)

country_factor_map_mean <- PTCA4CATA::createFactorMap(country_env_mca$TExPosition.Data$fi,
                                                 col.points = unique(country_env_mca$Plotting.Data$fi.col),
                                                 col.labels = unique(country_env_mca$Plotting.Data$fi.col),
                                                 axis1 = axis1,
                                                 axis2 = axis2,
                                                 display.labels = TRUE,
                                                 cex = 8,alpha.points = 0.8)

country_label4Map <- PTCA4CATA::createxyLabels.gen(axis1,axis2,lambda = country_env_mca$TExPosition.Data$eigs, tau = country_env_mca$TExPosition.Data$t) 



country_map = country_factor_map$zeMap + country_label4Map + country_factor_map_mean$zeMap_dots + country_factor_map_mean$zeMap_text


GraphElli <- PTCA4CATA::MakeCIEllipses(BootCube.Gr$BootCube[,c(axis1, axis2),],
                                       names.of.factors = c(paste("Dimension",axis1), paste("Dimension",axis2)),
                                       col = unique(country_env_mca$Plotting.Data$fi.col),
                                       p.level = .95
)


print(country_map + GraphElli)


```


**With Tolerance Interval**

```{r echo=FALSE}


  
axis1 = 1
axis2 = 2

country_factor_map <- PTCA4CATA::createFactorMap(country_env_mca$TExPosition.Data$fii, title='', 
                                                 col.points = country_env_mca$Plotting.Data$fii.col,
                                                 col.labels = country_env_mca$Plotting.Data$fii.col,
                                                 axis1 = axis1,
                                                 axis2 = axis2,
                                                 display.labels = FALSE)

country_factor_map_mean <- PTCA4CATA::createFactorMap(country_env_mca$TExPosition.Data$fi,
                                                 col.points = unique(country_env_mca$Plotting.Data$fi.col),
                                                 col.labels = unique(country_env_mca$Plotting.Data$fi.col),
                                                 axis1 = axis1,
                                                 axis2 = axis2,
                                                 display.labels = TRUE,
                                                 cex = 8,alpha.points = 0.8)

country_label4Map <- PTCA4CATA::createxyLabels.gen(axis1,axis2,lambda = country_env_mca$TExPosition.Data$eigs, tau = country_env_mca$TExPosition.Data$t) 



country_map = country_factor_map$zeMap + country_label4Map + country_factor_map_mean$zeMap_dots + country_factor_map_mean$zeMap_text


GraphTI.Hull <- PTCA4CATA::MakeToleranceIntervals(country_env_mca$TExPosition.Data$fii[,c(axis1, axis2)],
                                                  design = country_env_df$Happiness_Rank,
                                                  # line below is needed
                                                  names.of.factors =  c("Dim1","Dim2"), # needed 
                                                  col = unique(country_env_mca$Plotting.Data$fii.col),
                                                  line.size = .50, 
                                                  line.type = 3,
                                                  alpha.ellipse = .2,
                                                  alpha.line    = .4,
                                                  p.level       = .75)


print(country_map + GraphTI.Hull)




```



```{r echo=FALSE, results='hide'}
loop <- matrix(data = c(1,2, 7,9), nrow = 2, ncol = 2, byrow = TRUE)
color4Authors <- country_env_mca$Plotting.Data$fi.col
Fi <- country_env_mca$ExPosition.Data$fi
Fj <- country_env_mca$ExPosition.Data$fj
```

#### Loadings

```{r echo=FALSE}

col_palate = brewer.pal(n = 12, name='Set3')

col4J = vector('list', nrow(country_env_mca$TExPosition.Data$fj))
col4J[grep('rain',rownames(country_env_mca$TExPosition.Data$fj))] = col_palate[1]
col4J[grep('temp',rownames(country_env_mca$TExPosition.Data$fj))] = 'red' #col_palate[2]
col4J[sapply(col4J, 'is.null')] = rep(col_palate[3:11], length.out = sum(sapply(col4J, 'is.null')))

  
  axis1 = 1
  axis2 = 2

  country_factor_map <- PTCA4CATA::createFactorMap(country_env_mca$TExPosition.Data$fj, title='', 
                                                 col.points = col4J,
                                                 col.labels = col4J,
                                                 axis1 = axis1,
                                                 axis2 = axis2,
                                                 display.labels = TRUE)

country_label4Map <- PTCA4CATA::createxyLabels.gen(axis1,axis2,lambda = country_env_mca$TExPosition.Data$eigs, tau = country_env_mca$TExPosition.Data$t) 


#country_factor_mean_map <- PTCA4CATA::createFactorMap(axis_mean, title='', 
#                                                 col.points = col_mean,
#                                                 col.labels = col_mean,
#                                                 axis1 = axis1,
#                                                 axis2 = axis2,
#                                                 display.labels = TRUE)


print(country_factor_map$zeMap + country_label4Map)


```

#### Loadings (correlation plot)
```{r echo=FALSE}

country_env_df_for_mca_int <- dummy.data.frame(country_env_df_for_mca)  #data.frame(lapply(country_env_df_for_mca, as.integer)) # get dummy

loadings_2 <- cor(country_env_df_for_mca_int, country_env_mca$TExPosition.Data$fii)

col_palate = brewer.pal(n = 12, name='Set3')

col4J = vector('list', nrow(loadings_2))
col4J[grep('rain',rownames(country_env_mca$TExPosition.Data$fii))] = col_palate[1]
col4J[grep('temp',rownames(country_env_mca$TExPosition.Data$fii))] = 'red' #col_palate[2]
col4J[sapply(col4J, 'is.null')] = rep(col_palate[3:11], length.out = sum(sapply(col4J, 'is.null')))

  
  
axis1 = 1
axis2 = 2
    
  
  loadings_map <- PTCA4CATA::createFactorMap(loadings_2, 
                          col.points = col4J, 
                          col.labels = col4J, 
                          axis1=1,
                          axis2=2,
                          constraints = list(minx = -1, miny = -1, maxx = 1 , maxy = 1)) 
  
  country_label4Map <- PTCA4CATA::createxyLabels.gen(axis1,axis2,lambda = country_env_mca$TExPosition.Data$eigs, tau = country_env_mca$TExPosition.Data$t) 
  
  corr_map <- loadings_map$zeMap_background  +  country_label4Map + PTCA4CATA::addCircleOfCor() +
              loadings_map$zeMap_text + PTCA4CATA::addArrows(loadings_2)#, color = col4J) 
  
  print(corr_map)

```


#### Most Contributing Variables (Inference)

Let's plot variable contributions against each chosen components i.e. 1, 2.

* With Bootstrap Ratio

```{r }

BR <- country_env_mca_inf$Inference.Data$boot.data$fj.boot.data$tests$boot.ratios

for (i in c(1, 2)) {
  laDim = i
  ba001.BR1 <- PrettyBarPlot2(BR[,laDim],
                              threshold = 2,
                              font.size = 5,
                              color4bar = gplots::col2hex(col4J), # we need hex code
                              main = paste0('DiCA: Bootstrap ratio ',laDim),
                              ylab = 'Bootstrap ratios',horizontal = FALSE
                              #ylim = c(1.2*min(BR[,laDim]), 1.2*max(BR[,laDim]))
  )
  print(ba001.BR1)
}

```

#### Permutation Test

```{r echo=FALSE}
for (i in c(1, 2)) {
zeDim = i
pH1 <- prettyHist(
  distribution = country_env_mca_inf$Inference.Data$components$eigs.perm[,zeDim], 
  observed = country_env_mca_inf$Fixed.Data$TExPosition.Data$eigs[zeDim], 
  #xlim = c(country_env_mca_inf$Fixed.Data$TExPosition.Data$eigs[zeDim] - 2, country_env_mca_inf$Fixed.Data$TExPosition.Data$eigs[zeDim]+2), # needs to be set by hand
  xlim = c(-0.001, 0.01),
  breaks = 20,
  border = "black", 
  main = paste0("DiCA: Permutation Test for Eigenvalue ",zeDim),
  xlab = paste0("Eigenvalue ",zeDim), 
  ylab = "", 
  counts = FALSE, 
  cutoffs = c(0.025, 0.975))
}
```

#### Parallet Test

```{r echo=FALSE}
country_env_mca_mc <- data4PCCAR::monteCarlo.eigen(X = country_env_df_for_mca, nIter = 1000)
for (i in c(1, 2)) {
  zeDim = i
  pH1.p <- prettyHist(country_env_mca_mc$rand.eigs[,zeDim], 
                    observed = country_env_mca_mc$fixed.eigs[zeDim], 
                    xlim = c(0, country_env_mca_mc$fixed.eigs[zeDim]+2), # needs to set by hand
                    breaks = 20,
                    border = "black", 
                    main = paste0("DiCA - Monte Carlo (Parallel) Test for Eigenvalue ",zeDim),
                    xlab = paste0("Eigenvalue ",zeDim), 
                    ylab = "", 
                    counts = FALSE, 
                    cutoffs = c(0.025,0.975))
}

```

#### Bootstrap Test

```{r}

#country_env_pca_br <- PTCA4CATA::Boot4Mean(country_env_pca$ExPosition.Data$fi, design = country_env_df$Happiness_Rank, niter=100, suppressProgressBar = FALSE)
country_env_mca_bs <- data4PCCAR::boot.eigen(X = country_env_df_for_mca, nIter = 1000)

for (i in c(1, 2)) {
  zeDim = i
  prettyHist(country_env_mca_bs$boot.eigs[,zeDim], 
                    observed = country_env_mca_bs$fixed.eigs[zeDim], 
                    xlim = c(0, country_env_mca_bs$fixed.eigs[zeDim]+2), # needs to set by hand
                    breaks = 20,
                    border = "black", 
                    main = paste0("Bootstrapped distribution for Eigenvalue ",zeDim),
                    xlab = paste0("Eigenvalue ",zeDim), 
                    ylab = "", 
                    counts = FALSE, 
                    cutoffs = c(0.025, 0.975))
}
```



### Conclusion
Among PCA, MCA, BADA and DiCA, DiCA is able to best seperate the data based on levels of happiness. It seems increase in temperature is correlated with increase in happiness and increase in rain decreases happiness.

