---
title: "All Countries Environmental Data"
output:
  word_document: default
  html_notebook: default
  pdf_document: default
---


```{r include=FALSE, results='hide'}
# Installing required packages
# sudo apt-get install libcairo2-dev
# sudo apt-get install libxml2-dev

devtools::install_github('HerveAbdi/PTCA4CATA')
library(corrplot)
library(ExPosition)
library(InPosition)
library(animation)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(gplots)
library(PTCA4CATA)
library(purrr)
library(tidyr)
library(ExPosition)
library(InPosition)
library(dummies)
# use the last version of PTCA

# also install data4PCCAR last version
devtools::install_github('HerveAbdi/data4PCCAR')
library(data4PCCAR)
library(cowplot)
rm(list = ls())

```

# Multiple Component Analysis

## Description

Multiple correspondence analysis (MCA) is an extension of correspondence analysis(CA) which allows one to analyze the pattern of relationships of several categorical dependent variables. As such, it can also be seen as a generalization of principal component analysis when the variables to be analyzed are categorical instead of quantitative. Because MCA has been (re)discovered many times, equivalent methods are known under several different names such as optimal scaling, optimal or appropriate scoring, dual scaling, homogeneity analysis,scalogram analysis, and quantiﬁcation method.

**Interpreting MCA**
Multiple correspondence analysis locates all the categories in a Euclidean space.

* The first two dimensions of this space are plotted to examine the associations among the categories. 
* The top-right quadrant of the plot shows the categories.
* The bottom-left quadrant shows the association.
* This interpretation is based on points found in approximately the same direction from the origin and in approximately the same region of the space. Distances between points do not have a straightforward interpretation.


```{r echo=FALSE, results='hide'}
country_env_df = read.csv('../dataset/World_countries_env_vars.csv')
country_env_df_for_corr = country_env_df[, !names(country_env_df) %in% c('Country')]

country_env_df$Happiness_Rank = cut(country_env_df$Happiness_Rank,breaks = c(0, 39,118, 159), right=TRUE, labels = c('VH' , 'H', 'U'),ordered_result = TRUE)
```

## Density Plot
Let's observe the distribution of each variables to get an intuition of how we can bin these variables. It's important to have nearly equal number of observations in the each bin and to try to cut the variables in a way to so that each new binned distribution is nearly Gaussian. We can also verify that our binning is appropiate by calculating Spearman Correlation for each of original variable and binned variable, the correlation coefficient should be close to 1.

```{r echo=FALSE}

country_env_df_for_pca = country_env_df[, !names(country_env_df) %in% c('Country', 'Happiness_Rank')]

rownames(country_env_df_for_corr) <- paste0(country_env_df[,'Country'], '(' ,country_env_df[,'Happiness_Rank'], ')')
rownames(country_env_df_for_pca) <- paste0(country_env_df[,'Country'], '(' ,country_env_df[,'Happiness_Rank'], ')')

```


```{r echo=FALSE}

cut_points = list()

cut_points$accessibility_to_cities <- c(200,500)
cut_points$aspect <- c(160,180)
cut_points$cloudiness <- c(150,220)
cut_points$cropland_cover <- c(20,40)
cut_points$elevation <- c(200,800)
cut_points$isothermality <- c(40,65)
cut_points$rain_coldestQuart <- c(150, 300)
cut_points$rain_driestMonth <- c(15, 40)
cut_points$rain_driestQuart <- c(70,150)
cut_points$rain_mean_annual <- c(500,1500)
cut_points$rain_seasonailty <- c(50, 80)
cut_points$rain_warmestQuart <- c(100,300)
cut_points$rain_wettestMonth <- c(100, 300)
cut_points$rain_wettestQuart <- c(250, 700)
cut_points$slope <- c(0.7, 1.5)
cut_points$temp_annual_range <- c(20,30)
cut_points$temp_coldestQuart <- c(0,15)
cut_points$temp_diurnal_range <- c(10,14)
cut_points$temp_driestQuart <- c(10,20)
cut_points$temp_max_warmestMonth <- c(25,35)
cut_points$temp_mean_annual <- c(12,25)
cut_points$temp_min_coldestMonth <- c(0,10)
cut_points$temp_seasonality <- c(30, 70)
cut_points$temp_warmestQuart <- c(20, 25)
cut_points$temp_wettestQuart <- c(15, 25)
cut_points$tree_canopy_cover <- c(10, 30)
cut_points$wind <- c(2,3,4)
```


```{r echo=FALSE}

plot_bins <- function(df, col) {
  plot_col <- ggplot(df) + aes(df[,col]) +  geom_density() + 
    geom_vline(xintercept = cut_points[[col]], color = 'red') +
    xlab(col) + theme_minimal() + ylab(NULL) +
  theme(axis.text.y = element_blank())
  return(plot_col)
}

bin_hist_list <- lapply(colnames(country_env_df_for_pca), function (col) plot_bins(country_env_df_for_pca, col))

plot_grid(plotlist=bin_hist_list)

```


## Binning

Structure of Data after binning based on above observation.

```{r echo=FALSE}

cut_r <- function(x, b = 4 ) {
  c <- cut(x,breaks = b, include.lowest = TRUE, labels = FALSE)
  return (factor(c))
}

country_env_df_for_mca = mapply(function(col, col_name) cut_r(col, c(min(col), cut_points[[col_name]], max(col))) , country_env_df_for_pca, colnames(country_env_df_for_pca)  )
country_env_df_for_mca <- as.data.frame(country_env_df_for_mca)

str(country_env_df_for_mca)
```

## Spearman Correlation

Let's observe correlation between original data and binned data to make sure that neither the correlation ceofficient is too low or perfect.
```{r echo=FALSE}

cor_spear <- mapply(function(x,y) cor(x, as.integer(y),method = "spearman"), country_env_df_for_pca, country_env_df_for_mca)
#columns = colnames(country_env_df_for_pca)
#cor_df <- data.frame(col = columns, corr = cor_spear)
cor_p <- as.data.frame(cor_spear)

ggplot(data=cor_p, aes(x=rownames(cor_p), y=cor_p$cor_spear)) +
    geom_bar(stat="identity") + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) +
  xlab("") + ylab("Spearmen Correlation") + ylim(0, 1)
  

```

## Correlation Plot
Visually analyze multicollinearity in the system of the original data

```{r echo=FALSE}
corr_result = cor(country_env_df_for_corr)
corrplot(corr_result,order = 'hclust', addrect = 7)
```

## Heatmap

For binned data.

```{r echo=FALSE}

corrMatBurt.list <- phi2Mat4BurtTable(country_env_df_for_mca)
corr4MCA <- corrplot.mixed(as.matrix(corrMatBurt.list$phi2.mat,
                                     title = "Phi2: (squared) Correlation Map for MCA"))
a0001a.corMat.phi2 <- recordPlot()

```


```{r mca, echo=FALSE, results='hide', message=FALSE}

country_env_mca <- epMCA(DATA = country_env_df_for_mca, DESIGN = country_env_df$Happiness_Rank, graphs = FALSE,symmetric = TRUE)
country_env_mca_a <- epMCA(DATA = country_env_df_for_mca, DESIGN = country_env_df$Happiness_Rank, graphs = FALSE, symmetric = FALSE)

country_env_mca_inf <- InPosition::epMCA.inference.battery(DATA = country_env_df_for_mca, DESIGN = country_env_df$Happiness_Rank, graphs = FALSE)


col_palate = brewer.pal(n = 12, name='Set3')
  
col4J = vector('list', nrow(country_env_mca$ExPosition.Data$fj))
col4J[grep('rain',rownames(country_env_mca$ExPosition.Data$fj))] = col_palate[1]
col4J[grep('temp',rownames(country_env_mca$ExPosition.Data$fj))] = 'red' #col_palate[2]
col4J[sapply(col4J, 'is.null')] = rep( col_palate[c(-1)], length.out = 28)

#length(col4J)
#col4J[sapply(col4J, 'is.null')]
```

## Scree Plot
Gives amount of information explained by corresponding component. Gives an intuition to decide which components best represent data in order to answer the research question.

P.S. The most contribution component may not always be most useful for a given research question. 


```{r echo=FALSE}

PTCA4CATA::PlotScree(ev = country_env_mca$ExPosition.Data$eigs,
                      p.ev =  country_env_mca_inf$Inference.Data$components$p.vals,
                      title = 'SCREE Plot',
                      plotKaiser = TRUE
)

```

## Factor Scores 

Lets visualize happiness categories for components 1-10, to make a decision (visually) on the most important components.


```{r echo=FALSE}

loop <- matrix(data = c(1,2, 7,9), nrow = 2, ncol = 2, byrow = TRUE)

list_color <- c('#305ABF' = 'burlywood3', '#84BF30' = 'royalblue1', '#BF30AD' = 'springgreen4')
happy_levels <- c('Happy', 'Happier', 'Happiest')

country_env_mca$Plotting.Data$fi.col[,1] <- recode (country_env_mca$Plotting.Data$fi.col, !!!list_color)

groups_of_data <- vector('list', 3)

col = country_env_mca$Plotting.Data$fi.col
fi = country_env_mca$ExPosition.Data$fi
j = 1
color_filter = col == unique(col)[1]
groups_of_data[[j]] = color_filter
j = j + 1

color_filter = col == unique(col)[2]
groups_of_data[[j]] = color_filter
j = j + 1

color_filter = col == unique(col)[3]
groups_of_data[[j]] = color_filter
j = j + 1

country_env_mca_mean = aggregate(x=country_env_mca$ExPosition.Data$fi, by = split(col, col(col)), FUN = mean)
rownames(country_env_mca_mean) <- happy_levels

country_env_mca_mean <- country_env_mca_mean[-1]

```

**With Confidence Interval**
```{r echo=FALSE}
BootCube.Gr <- PTCA4CATA::Boot4Mean(country_env_mca$ExPosition.Data$fi, 
                                    design = country_env_df$Happiness_Rank,
                                    niter = 1000,
                                    suppressProgressBar = TRUE)
for (i in 1:2){
  
  axis1 = loop[i,1]
  axis2 = loop[i,2]

  country_factor_map <- PTCA4CATA::createFactorMap(country_env_mca$ExPosition.Data$fi, title='', 
                                                 col.points = country_env_mca$Plotting.Data$fi.col,
                                                 col.labels = country_env_mca$Plotting.Data$fi.col,
                                                 axis1 = axis1,
                                                 axis2 = axis2,
                                                 display.labels = FALSE)

country_factor_map_mean <- PTCA4CATA::createFactorMap(country_env_mca_mean,
                                                 col.points = unique(country_env_mca$Plotting.Data$fi.col),
                                                 col.labels = unique(country_env_mca$Plotting.Data$fi.col),
                                                 axis1 = axis1,
                                                 axis2 = axis2,
                                                 display.labels = TRUE,
                                                 cex = 8,alpha.points = 0.8)

country_label4Map <- PTCA4CATA::createxyLabels.gen(axis1,axis2,lambda = country_env_mca$ExPosition.Data$eigs, tau = country_env_mca$ExPosition.Data$t) 



country_map = country_factor_map$zeMap + country_label4Map + country_factor_map_mean$zeMap_dots + country_factor_map_mean$zeMap_text


GraphElli <- PTCA4CATA::MakeCIEllipses(BootCube.Gr$BootCube[,c(axis1, axis2),],
                                       names.of.factors = c(paste("Dimension",axis1), paste("Dimension",axis2)),
                                       col = unique(country_env_mca$Plotting.Data$fi.col),
                                       p.level = .95
)

country_map = country_map + GraphElli

print(country_map)
}

```


**With Tolerance Interval**

```{r echo=FALSE}

for (i in 1:2){
  
  axis1 = loop[i,1]
  axis2 = loop[i,2]

country_factor_map <- PTCA4CATA::createFactorMap(country_env_mca$ExPosition.Data$fi, title='', 
                                                 col.points = country_env_mca$Plotting.Data$fi.col,
                                                 col.labels = country_env_mca$Plotting.Data$fi.col,
                                                 axis1 = axis1,
                                                 axis2 = axis2,
                                                 display.labels = FALSE)

country_factor_map_mean <- PTCA4CATA::createFactorMap(country_env_mca_mean,
                                                 col.points = unique(country_env_mca$Plotting.Data$fi.col),
                                                 col.labels = unique(country_env_mca$Plotting.Data$fi.col),
                                                 axis1 = axis1,
                                                 axis2 = axis2,
                                                 display.labels = TRUE,
                                                 cex = 8,alpha.points = 0.8)

country_label4Map <- PTCA4CATA::createxyLabels.gen(axis1,axis2,lambda = country_env_mca$ExPosition.Data$eigs, tau = country_env_mca$ExPosition.Data$t) 



country_map = country_factor_map$zeMap + country_label4Map + country_factor_map_mean$zeMap_dots + country_factor_map_mean$zeMap_text


GraphTI.Hull <- PTCA4CATA::MakeToleranceIntervals(country_env_mca$ExPosition.Data$fi[,c(axis1, axis2)],
                                                  design = country_env_df$Happiness_Rank,
                                                  # line below is needed
                                                  names.of.factors =  c("Dim1","Dim2"), # needed 
                                                  col = unique(country_env_mca$Plotting.Data$fi.col),
                                                  line.size = .50, 
                                                  line.type = 3,
                                                  alpha.ellipse = .2,
                                                  alpha.line    = .4,
                                                  p.level       = .75)

country_map = country_map + GraphTI.Hull

print(country_map)


}

```



```{r echo=FALSE, results='hide'}
loop <- matrix(data = c(1,2, 7,9), nrow = 2, ncol = 2, byrow = TRUE)
color4Authors <- country_env_mca$Plotting.Data$fi.col
Fi <- country_env_mca$ExPosition.Data$fi
Fj <- country_env_mca$ExPosition.Data$fj
Fj.a <- country_env_mca_a$ExPosition.Data$fj
```

## Loadings

```{r echo=FALSE}

col_palate = brewer.pal(n = 12, name='Set3')

col4J = vector('list', nrow(country_env_mca$ExPosition.Data$fj))
col4J[grep('rain',rownames(country_env_mca$ExPosition.Data$fj))] = col_palate[1]
col4J[grep('temp',rownames(country_env_mca$ExPosition.Data$fj))] = 'red' #col_palate[2]
col4J[sapply(col4J, 'is.null')] = col_palate[3] #rep(col_palate[3:11], length.out = sum(sapply(col4J, 'is.null')))

#axis_mean = aggregate(country_env_mca$ExPosition.Data$fj, list(unlist(col4J)), mean)
#col4mean = axis_mean[,1]
#rownames(axis_mean) = c('Others', 'Rain', 'Temp')
#axis_mean = axis_mean[,2:ncol(axis_mean)]

for (i in 1:2){
  
  axis1 = loop[i,1]
  axis2 = loop[i,2]

  country_factor_map <- PTCA4CATA::createFactorMap(country_env_mca$ExPosition.Data$fj, title='', 
                                                 col.points = col4J,
                                                 col.labels = col4J,
                                                 axis1 = axis1,
                                                 axis2 = axis2,
                                                 display.labels = TRUE,alpha.labels = 0.4, 
                                                 text.cex = 2.2)

country_label4Map <- PTCA4CATA::createxyLabels.gen(axis1,axis2,lambda = country_env_mca$ExPosition.Data$eigs, tau = country_env_mca$ExPosition.Data$t) 

#country_factor_mean_map <- PTCA4CATA::createFactorMap(axis_mean, title='', 
#                                                 col.points = col4mean,
#                                                 col.labels = col4mean,
#                                                 axis1 = axis1,
#                                                 axis2 = axis2,
#                                                 display.labels = TRUE,cex = 4)

country_map = country_factor_map$zeMap + country_label4Map 
#+ country_factor_mean_map$zeMap_dots + country_factor_mean_map$zeMap_text

print(country_map)
}

```

## Loadings (correlation plot)
```{r echo=FALSE}

country_env_df_for_mca_int <- dummy.data.frame(country_env_df_for_mca)  #data.frame(lapply(country_env_df_for_mca, as.integer)) # get dummy

loadings_2 <- cor(country_env_df_for_mca_int, country_env_mca$ExPosition.Data$fi)

col_palate = brewer.pal(n = 12, name='Set3')

col4J = vector('list', nrow(loadings_2))
col4J[grep('rain',rownames(country_env_mca$ExPosition.Data$fj))] = col_palate[1]
col4J[grep('temp',rownames(country_env_mca$ExPosition.Data$fj))] = 'red' #col_palate[2]
col4J[sapply(col4J, 'is.null')] = rep(col_palate[3:11], length.out = sum(sapply(col4J, 'is.null')))

  

for (i in 1:2){
  
  axis1 = loop[i,1]
  axis2 = loop[i,2]

    
  
  loadings_map <- PTCA4CATA::createFactorMap(loadings_2, 
                          col.points = col4J, 
                          col.labels = col4J, 
                          axis1=1,
                          axis2=2,alpha.labels = 0.4, 
                          text.cex = 2.2,
                          constraints = list(minx = -1, miny = -1, maxx = 1 , maxy = 1)) 
  
  country_label4Map <- PTCA4CATA::createxyLabels.gen(axis1,axis2,lambda = country_env_mca$ExPosition.Data$eigs, tau = country_env_mca$ExPosition.Data$t) 
  
  corr_map <- loadings_map$zeMap_background  +  country_label4Map + PTCA4CATA::addCircleOfCor() +
              loadings_map$zeMap_text + PTCA4CATA::addArrows(loadings_2, color = col4J) 
  
  print(corr_map)
}
```


## Most Contributing Variables (Inference)

Let's plot variable contributions against each chosen components i.e. 1, 2, 7, 9.

* With Bootstrap Ratio

```{r echo=FALSE}

BR <- country_env_mca_inf$Inference.Data$fj.boots$tests$boot.ratios

for (i in c(1, 2, 7, 9)) {
  laDim = i
  ba001.BR1 <- PrettyBarPlot2(BR[,laDim],
                              threshold = 2,
                              font.size = 5,
                              color4bar = gplots::col2hex(col4J), # we need hex code
                              main = paste0('Bootstrap ratio ',laDim),
                              ylab = 'Bootstrap ratios',horizontal = FALSE
                              #ylim = c(1.2*min(BR[,laDim]), 1.2*max(BR[,laDim]))
  )
  print(ba001.BR1)
}

```

## Permutation Test

```{r echo=FALSE}
for (i in c(1, 2, 7, 9)) {
zeDim = i
xmin = min(country_env_mca_inf$Inference.Data$components$eigs.perm[,zeDim] - mean(country_env_mca_inf$Inference.Data$components$eigs.perm[,zeDim]))
xmax = max(country_env_mca_inf$Inference.Data$components$eigs.perm[,zeDim] + mean(country_env_mca_inf$Inference.Data$components$eigs.perm[,zeDim]))
pH1 <- prettyHist(
  distribution = country_env_mca_inf$Inference.Data$components$eigs.perm[,zeDim], 
  observed = country_env_mca_inf$Fixed.Data$ExPosition.Data$eigs[zeDim], 
  xlim = c(-xmin, xmax), # needs to be set by hand
  breaks = 20,
  border = "black", 
  main = paste0("MCA: Permutation Test for Eigenvalue ",zeDim),
  xlab = paste0("Eigenvalue ",zeDim), 
  ylab = "", 
  counts = FALSE, 
  cutoffs = c( 0.975))
}
```

## Parallet Test

```{r echo=FALSE}

country_env_df_for_mca_int <- lapply(country_env_df_for_mca, function(x) as.numeric(as.character(x)))

country_env_df_for_mca_int = as.data.frame(country_env_df_for_mca_int)

country_env_mca_mc <- data4PCCAR::monteCarlo.eigen(X = country_env_df_for_mca_int, nIter = 1000)
for (i in c(1, 2, 7, 9)) {
  zeDim = i
  pH1.p <- prettyHist(country_env_mca_mc$rand.eigs[,zeDim], 
                    observed = country_env_mca_mc$fixed.eigs[zeDim], 
                    xlim = c(0, country_env_mca_mc$fixed.eigs[zeDim]+2), # needs to set by hand
                    breaks = 20,
                    border = "black", 
                    main = paste0("MCA: Monte Carlo (Parallel) Test for Eigenvalue ",zeDim),
                    xlab = paste0("Eigenvalue ",zeDim), 
                    ylab = "", 
                    counts = FALSE, 
                    cutoffs = c( 0.975))
}

```

## Bootstrap Test

```{r echo=FALSE}

#country_env_pca_br <- PTCA4CATA::Boot4Mean(country_env_pca$ExPosition.Data$fi, design = country_env_df$Happiness_Rank, niter=100, suppressProgressBar = FALSE)
country_env_mca_bs <- data4PCCAR::boot.eigen(X = country_env_df_for_mca_int, nIter = 1000)

for (i in c(1, 2, 7, 9)) {
  zeDim = i
  prettyHist(country_env_mca_bs$boot.eigs[,zeDim], 
                    observed = country_env_mca_bs$fixed.eigs[zeDim], 
                    xlim = c(0, country_env_mca_bs$fixed.eigs[zeDim]+2), # needs to set by hand
                    breaks = 20,
                    border = "black", 
                    main = paste0("MCA: Bootstrapped distribution for Eigenvalue ",zeDim),
                    xlab = paste0("Eigenvalue ",zeDim), 
                    ylab = "", 
                    counts = FALSE, 
                    cutoffs = c(0.025, 0.975))
}
```

## Conclusion
* Component 1: 
  + Rows: Normal & Happy
  + Columns: Cloudiness & Rain vs Cropland, Aspect, Elevation
  + Interpret: People in countries with more Cloudiness, Trees and Rain tends to be happier.

* Component 7: 
  + Rows: Happy & Unhappy
  + Columns: Temp and Rain vs Accessibility and Cropland
  + Interpret: Rain and Temp seems to be main reason for unhappiness and Cropland is important for Happiness.

* Component 9: 
  + Rows: Happy & Very Happy
  + Columns: Temp vs Rain
  + Interpret: Rain and Temp seems to be main reason for Happiness. *This contradicts with Component 7 and 1*.

