---
title: "Ritesh_Malaiya_Group_1_DiSTATIS"
output:
  html_document: default
  github_document: default
  word_document: default
---


```{r cars, echo=FALSE}

rm(list = ls())

devtools::install_github('HerveAbdi/PTCA4CATA')
suppressMessages(library(PTCA4CATA))
# PTCA4CATA should first to avoid conflict with TInPosition
suppressMessages(library(ExPosition))
#suppressMessages(library(InPosition))
#suppressMessages(library(TInPosition))
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
suppressMessages(library(officer))
suppressMessages(library(flextable))
suppressMessages(library(rvg))
suppressMessages(library(useful))
suppressMessages(library(DistatisR))
library(RColorBrewer)


# Source the function file:
#
# install.packages('TExPosition')
# install.packages('MExPosition')
library(MExPosition)
rm(list = ls())
```

# DiSTATIS

## Description

DISTATIS is a new method that can be used to compare algorithms when their outputs consist of distance matrices computed on the same set of objects. The method ﬁrst evaluates the similarity between algorithms using a coefﬁcient called the RV coefﬁcient. From this analysis, a compromise matrix is computed which represents the best aggregate of the original matrices. In order to evaluate the differences between algorithms, the original distance matrices are then projected onto the compromise. The goal of DISTATIS is to analyze a set of distance matrices. In order to compare distance matrices, DISTATIS combines them into a common structure called a compromise and then projects the original distance matrices onto this compromise. 

```{r include=FALSE, results='hide'}
country_env_df = read.csv('../dataset/World_countries_env_vars.csv')
country_env_df_for_corr = country_env_df[, !names(country_env_df) %in% c('Country')]

country_env_df$Happiness_Rank = cut(country_env_df$Happiness_Rank,breaks = c(0, 39,118, 159), right=TRUE, labels = c('VH' , 'H', 'U'),ordered_result = TRUE)

str(country_env_df)
```


```{r echo=FALSE}

country_env_df_for_mfa = country_env_df[, !names(country_env_df) %in% c('Country', 'Happiness_Rank')]

rownames(country_env_df_for_corr) <- paste0(country_env_df[,'Country'], '(' ,country_env_df[,'Happiness_Rank'], ')')
rownames(country_env_df_for_mfa) <- paste0(country_env_df[,'Country'], '(' ,country_env_df[,'Happiness_Rank'], ')')

column.design = rep('Others', length.out = ncol(country_env_df_for_mfa) )

column.design[ grep('rain',colnames(country_env_df_for_mfa)) ] = 'Rain'
column.design[ grep('temp',colnames(country_env_df_for_mfa)) ] = 'Temp'
column.design = t(as.matrix(column.design))
colnames(column.design) = colnames(country_env_df_for_mfa)
rownames(column.design) = 'Type'
```

```{r echo=FALSE}
#Raw_Data <- read.csv('../dataset/natural36_constrained_forR.csv', row.names = 1)



Sorting_Data <- country_env_df_for_mfa #Raw_Data[-37,]
Design_Data <- column.design #Raw_Data[37,]

Design_cat <- column.design #country_env_df$Happiness_Rank 
#cut(t(Design_Data[1,]), breaks = c(0,1,4,15), include.lowest=TRUE, right = TRUE, labels = c('Non-Musician', 'Moderate', 'Musician'))

```

```{r nice-tab-distatis, tidy=FALSE, echo=FALSE}
#knitr::kable(
#  head(country_env_df_for_mfa[,1:10]), caption = '4 Pianist for each of 3 Composers',
#  booktabs = TRUE
#)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
DistanceCube <- DistatisR::DistanceFromSort(Sorting_Data)
testDistatis <- DistatisR::distatis(DistanceCube)
BootF <- BootFactorScores(testDistatis$res4Splus$PartialF)
```

## SCREE Plot - RV-MAT

```{r echo=FALSE}
ev4C <- testDistatis$res4Cmat$eigValues
Scree.1 <-  PlotScree(ev = ev4C,
          p.ev = NULL, max.ev = NULL, alpha = 0.05,
          col.ns = "#006D2C", col.sig = "#54278F", plotKaiser = TRUE,
          title = "RV-mat: Explained Variance per Dimension")

```

## Plotting Assessor Matrix

```{r echo=FALSE}

G <- testDistatis$res4Cmat$G
# Create a color scheme for the judges
col4B <- 'turquoise3'
col4G <- 'thistle3'

col_palate = brewer.pal(n = 12, name='Set3')

col4J = vector('list', nrow(G))
col4J[grep('rain',rownames(G))] = col_palate[1]
col4J[grep('temp',rownames(G))] = 'red' #col_palate[2]
col4J[sapply(col4J, 'is.null')] = col_palate[3] #rep(col_palate[4], length.out = sum(sapply(col4J, 'is.null')))

#col <- brewer.pal(n = 3, name='Set1')
#col4Musicians <- dplyr::recode(Design_cat, `Non-Musician` = col[1], `Moderate` = col[2], `Musician` = col[3])

baseMap.j <- PTCA4CATA::createFactorMap(G,
                                        title = 'The Rv map',
                                        col.points   = unlist(col4J), #col4Musicians,
                                        alpha.points =  .3,
                                        col.labels = unlist(col4J))
# A graph for the J-set
aggMap.j <- baseMap.j$zeMap_background + # background layer
           baseMap.j$zeMap_dots + baseMap.j$zeMap_text # dots & labels
# We print this Map with the following code
print(aggMap.j)

```



### ConvexHull

```{r echo=FALSE}

GraphTJ.Hull.100 <- MakeToleranceIntervals(G,
                                  as.factor(Design_cat),
                                  names.of.factors = c("Dim1","Dim2"),
                                  col = unique(col4J),
                                          line.size = .5,
                                          line.type = 3,
                                          alpha.ellipse = .1,
                                          alpha.line = .4,
                                          p.level = 1, # full Hulls
                                          type = 'hull' #
                                          # use 'hull' for convex hull
)

aggMap.j.withHull <- baseMap.j$zeMap_background + # background layer
                           baseMap.j$zeMap_dots + GraphTJ.Hull.100

JudgesMeans.tmp <- aggregate(G, list(as.factor(Design_cat)), mean) # compute the means
JudgesMeans <- JudgesMeans.tmp[,2:ncol(JudgesMeans.tmp )] # drop var 1

rownames(JudgesMeans) <- JudgesMeans.tmp[,1] # use var 1 to name the groups


col4Means <- unique(col4J)

MapGroup    <- PTCA4CATA::createFactorMap(JudgesMeans,
                                  axis1 = 1, axis2 = 2,
                                  constraints = baseMap.j$constraints,
                                          title = NULL,
                                          col.points = col4Means,
                                          display.points = TRUE,
                                          pch = 19, cex = 5,
                                          display.labels = TRUE,
                                          col.labels = col4Means,
                                          text.cex = 4,
                                          font.face = "bold",
                                          font.family = "sans",
                                          col.axes = "darkorchid",
                                          alpha.axes = 0.2,
                                          width.axes = 1.1,
                          col.background = adjustcolor("lavender",
                                             alpha.f = 0.2),
                                          force = 1, segment.size = 0)

aggMap.j.withMeans <- aggMap.j.withHull +
                    MapGroup$zeMap_dots + MapGroup$zeMap_text

print(aggMap.j.withMeans)

```

## SCREE Plot - SV-MAT

```{r echo=FALSE}

ev4S <- eigen(testDistatis$res4Splus$Splus,
                symmetric = TRUE, only.values = TRUE)$values
# A scree for the compromise
Scree.S <-  PlotScree(ev = ev4S,
                   p.ev = NULL, max.ev = NULL, alpha = 0.05,
                   col.ns = "#006D2C", col.sig = "#54278F",plotKaiser = TRUE,
                   title = "S-mat: Explained Variance per Dimension")

```

## I Set

```{r echo=FALSE}


Fi <- testDistatis$res4Splus$F
col <- brewer.pal(n = 3, name='Set2')

list_color <- c('U' = 'burlywood3', 'H' = 'royalblue1', 'VH' = 'springgreen4')

lx_col <- dplyr::recode (country_env_df$Happiness_Rank, `U` = 'burlywood3', `H` = 'royalblue1', `VH` = 'springgreen4')


#col4Composer <- prettyGraphsColorSelection(nrow(Fi))
#  Use colors from prettyGraphs
#---------------------------------------------------------------------
# Graphs for the I set
#---------------------------------------------------------------------
# Create the base map
constraints4Fi <- lapply(minmaxHelper(Fi),'*',1.2)
baseMap.i <- PTCA4CATA::createFactorMap(Fi,
                                        col.points = lx_col,
                                        col.labels = lx_col,
                        constraints = constraints4Fi,
                                        alpha.points =  .4)
#---------------------------------------------------------------------
# We are  interested about the labels here
# so we will use dots and labels
#---------------------------------------------------------------------
# Plain map with color for the I-set
aggMap.i <- baseMap.i$zeMap_background + baseMap.i$zeMap_dots +
                                         baseMap.i$zeMap_text
#---------------------------------------------------------------------
# print this Map
print(aggMap.i)

```


```{r echo=FALSE}

constraints4Fi <- lapply(minmaxHelper(Fi),'*',1.2)
GraphElli <- MakeCIEllipses(BootF[,1:2,],
                      names.of.factors = c("Factor 1","Factor 2"),
                      alpha.line = .5,
                      alpha.ellipse = .3,
                      line.size = .5,
                      line.type = 3,
                      col = lx_col,
                      p.level = .95 )
#---------------------------------------------------------------------
# create the I-map with Observations and their confidence intervals
#
aggMap.i.withCI <-  aggMap.i +  GraphElli + MapGroup$zeMap_text
#---------------------------------------------------------------------
# plot it!
print(aggMap.i.withCI)


```

## Cluster Analysis (Hartigan's Rule)
```{r echo=FALSE}
# Optimal number of clustera
#  install.packages('useful')
best.beers <- useful::FitKMeans(Fi, max.clusters = 5,
                                seed = 314)
print(best.beers) # when Hartigan parameter > 10 => add a cluster

```

```{r echo=FALSE}

plot.harti <- useful::PlotHartigan(best.beers)
print(plot.harti)

```


## Cluster Analysis (K-Means)

### Two centers
```{r echo=FALSE}




composer.kMeans <- kmeans(x = Fi , centers = 2)
col4Clusters  <- createColorVectorsByDesign(
              makeNominalData(
              as.data.frame(composer.kMeans$cluster)  ))


#col4Clus_unique <- unique(col4Clusters$oc)

#col <- brewer.pal(n = nrow(col4Clus_unique), name='Set1')

#col4Clusters$oc[ grep(col4Clus_unique[1],col4Clusters)] = col[1]
#col4Clusters$oc[ grep(col4Clus_unique[2],col4Clusters)] = col[2]
#col4Clusters$oc[ grep(col4Clus_unique[3],col4Clusters)] = col[3]

baseMap.i.km <- PTCA4CATA::createFactorMap(Fi,
                                        col.points = col4Clusters$oc,
                                        col.labels = col4Clusters$oc,
                                        constraints = constraints4Fi,
                                        alpha.points =  .3, alpha.labels = 0.4)

#---------------------------------------------------------------------
# Now to get a map by cluster:
col4Clusters  <- createColorVectorsByDesign(
              makeNominalData(
              as.data.frame(composer.kMeans$cluster)  ))

aggMap.i.km <- baseMap.i.km$zeMap_background +
  baseMap.i.km$zeMap_dots + baseMap.i.km$zeMap_text

col4C <- col4Clusters$gc[sort(rownames(col4Clusters$gc),
                                     index.return = TRUE)$ix]
#col4C <- col4Clus_unique
# create the map for the means
map4Clusters    <- PTCA4CATA::createFactorMap(composer.kMeans$centers,
                                          axis1 = 1, axis2 = 2,
                                          constraints = constraints4Fi,
                                          title = NULL,
                                          col.points = col4C,
                                          display.points = TRUE,
                                          pch = 19, cex = 5,
                                          display.labels = TRUE,
                                          col.labels = col4C,
                                          text.cex = 6,
                                          font.face = "bold",
                                          font.family = "sans",
                                          col.axes = "darkorchid",
                                          alpha.axes = 0.2,
                                          width.axes = 1.1,
                                          col.background =
                          adjustcolor("lavender", alpha.f = 0.2),
                                          force = 1, segment.size = 0)
# The map with observations and group means
aggMap.i.withCenters <- aggMap.i.km  +
               map4Clusters$zeMap_dots + map4Clusters$zeMap_text
#
print(aggMap.i.withCenters)

```


### Four centers
```{r echo=FALSE}




composer.kMeans <- kmeans(x = Fi , centers = 4)
col4Clusters  <- createColorVectorsByDesign(
              makeNominalData(
              as.data.frame(composer.kMeans$cluster)  ))


#col4Clus_unique <- unique(col4Clusters$oc)

#col <- brewer.pal(n = nrow(col4Clus_unique), name='Set1')

#col4Clusters$oc[ grep(col4Clus_unique[1],col4Clusters)] = col[1]
#col4Clusters$oc[ grep(col4Clus_unique[2],col4Clusters)] = col[2]
#col4Clusters$oc[ grep(col4Clus_unique[3],col4Clusters)] = col[3]

baseMap.i.km <- PTCA4CATA::createFactorMap(Fi,
                                        col.points = col4Clusters$oc,
                                        col.labels = col4Clusters$oc,
                                        constraints = constraints4Fi,
                                        alpha.points =  .3,alpha.labels = 0.4)

#---------------------------------------------------------------------
# Now to get a map by cluster:
col4Clusters  <- createColorVectorsByDesign(
              makeNominalData(
              as.data.frame(composer.kMeans$cluster)  ))

aggMap.i.km <- baseMap.i.km$zeMap_background +
  baseMap.i.km$zeMap_dots + baseMap.i.km$zeMap_text

col4C <- col4Clusters$gc[sort(rownames(col4Clusters$gc),
                                     index.return = TRUE)$ix]
#col4C <- col4Clus_unique
# create the map for the means
map4Clusters    <- PTCA4CATA::createFactorMap(composer.kMeans$centers,
                                          axis1 = 1, axis2 = 2,
                                          constraints = constraints4Fi,
                                          title = NULL,
                                          col.points = col4C,
                                          display.points = TRUE,
                                          pch = 19, cex = 5,
                                          display.labels = TRUE,
                                          col.labels = col4C,
                                          text.cex = 8,
                                          font.face = "bold",
                                          font.family = "sans",
                                          col.axes = "darkorchid",
                                          alpha.axes = 0.2,
                                          width.axes = 1.1,
                                          col.background =
                          adjustcolor("lavender", alpha.f = 0.2),
                                          force = 1, segment.size = 0)
# The map with observations and group means
aggMap.i.withCenters <- aggMap.i.km  +
               map4Clusters$zeMap_dots + map4Clusters$zeMap_text
#
print(aggMap.i.withCenters)

```



## Cluster Analysis (hclust)

### Two Clusters
```{r echo=FALSE}

beer.hc <- hclust(d = dist(Fi),
                  method = 'ward.D2' )

plot.tree <- plot(beer.hc,  main = "Beers. Ward's method")
#print(plot.tree)

#---------------------------------------------------------------------
# Use three clusters too
hc.3.cl <- rect.hclust(beer.hc, k = 2,
               border = c('darkorchid',
                          'darkolivegreen4','darkgoldenrod3')
                          )
#print(hc.3.cl)
```

### Fours Clusters

```{r echo=FALSE}

beer.hc <- hclust(d = dist(Fi),
                  method = 'ward.D2' )

plot.tree <- plot(beer.hc,  main = "Beers. Ward's method")
#print(plot.tree)

#---------------------------------------------------------------------
# Use three clusters too
hc.3.cl <- rect.hclust(beer.hc, k = 4,
               border = c('darkorchid',
                          'darkolivegreen4','darkgoldenrod3')
                          )
#print(hc.3.cl)
```

