---
title: "Ritesh_Malaiya_Group_1_DiSTATIS"
output: html_document
---


```{r cars}

rm(list = ls())

devtools::install_github('HerveAbdi/PTCA4CATA')
suppressMessages(library(PTCA4CATA))
# PTCA4CATA should first to avoid conflict with TInPosition
suppressMessages(library(ExPosition))
#suppressMessages(library(InPosition))
#suppressMessages(library(TInPosition))
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
suppressMessages(library(officer))
suppressMessages(library(flextable))
suppressMessages(library(rvg))
suppressMessages(library(useful))
suppressMessages(library(DistatisR))
library(RColorBrewer)


# Source the function file:
#
# install.packages('TExPosition')
# install.packages('MExPosition')
library(MExPosition)
rm(list = ls())
```

## DiSTATIS

### Dataset
```{r echo=FALSE}
Raw_Data <- read.csv('../dataset/natural36_constrained_forR.csv', row.names = 1)



Sorting_Data <- Raw_Data[-37,]
Design_Data <- Raw_Data[37,]

Design_cat <- cut(t(Design_Data[1,]), breaks = c(0,1,4,15), include.lowest=TRUE, right = TRUE, labels = c('Non-Musician', 'Moderate', 'Musician'))

```

```{r nice-tab-distatis, tidy=FALSE, echo=FALSE}
knitr::kable(
  head(Raw_Data, 20), caption = '4 Pianist for each of 3 Composers',
  booktabs = TRUE
)
```
```

```{r echo=FALSE}
DistanceCube <- DistatisR::DistanceFromSort(Sorting_Data)
testDistatis <- DistatisR::distatis(DistanceCube)
BootF <- BootFactorScores(testDistatis$res4Splus$PartialF)
```

### SCREE Plot

```{r echo=FALSE}
ev4C <- testDistatis$res4Cmat$eigValues
Scree.1 <-  PlotScree(ev = ev4C,
          p.ev = NULL, max.ev = NULL, alpha = 0.05,
          col.ns = "#006D2C", col.sig = "#54278F", plotKaiser = TRUE,
          title = "RV-mat: Explained Variance per Dimension")

```

### Plotting Assessor Matrix

```{r echo=FALSE}

G <- testDistatis$res4Cmat$G
# Create a color scheme for the judges
col4B <- 'turquoise3'
col4G <- 'thistle3'
col <- brewer.pal(n = 3, name='Set1')


col4Musicians <- dplyr::recode(Design_cat, `Non-Musician` = col[1], `Moderate` = col[2], `Musician` = col[3])

baseMap.j <- PTCA4CATA::createFactorMap(G,
                                        title = 'The Rv map',
                                        col.points   = col4Musicians,
                                        alpha.points =  .3,
                                        col.labels = col4Musicians)
# A graph for the J-set
aggMap.j <- baseMap.j$zeMap_background + # background layer
           baseMap.j$zeMap_dots + baseMap.j$zeMap_text # dots & labels
# We print this Map with the following code
print(aggMap.j)

```



### ConvexHull

```{r echo=FALSE}

GraphTJ.Hull.100 <- MakeToleranceIntervals(G,
                                  as.factor(Design_cat),
                                  names.of.factors = c("Dim1","Dim2"),
                                  col = unique(col4Musicians),
                                          line.size = .5,
                                          line.type = 3,
                                          alpha.ellipse = .1,
                                          alpha.line = .4,
                                          p.level = 1, # full Hulls
                                          type = 'hull' #
                                          # use 'hull' for convex hull
)

aggMap.j.withHull <- baseMap.j$zeMap_background + # background layer
                           baseMap.j$zeMap_dots + GraphTJ.Hull.100

JudgesMeans.tmp <- aggregate(G, list(Design_cat), mean) # compute the means
JudgesMeans <- JudgesMeans.tmp[,2:ncol(JudgesMeans.tmp )] # drop var 1

rownames(JudgesMeans) <- JudgesMeans.tmp[,1] # use var 1 to name the groups


col4Means <- unique(col4Musicians)

MapGroup    <- PTCA4CATA::createFactorMap(JudgesMeans,
                                  axis1 = 1, axis2 = 2,
                                  constraints = baseMap.j$constraints,
                                          title = NULL,
                                          col.points = col4Means,
                                          display.points = TRUE,
                                          pch = 19, cex = 5,
                                          display.labels = TRUE,
                                          col.labels = col4Means,
                                          text.cex = 4,
                                          font.face = "bold",
                                          font.family = "sans",
                                          col.axes = "darkorchid",
                                          alpha.axes = 0.2,
                                          width.axes = 1.1,
                          col.background = adjustcolor("lavender",
                                             alpha.f = 0.2),
                                          force = 1, segment.size = 0)

aggMap.j.withMeans <- aggMap.j.withHull +
                    MapGroup$zeMap_dots + MapGroup$zeMap_text

print(aggMap.j.withMeans)

```



```{r echo=FALSE}

ev4S <- eigen(testDistatis$res4Splus$Splus,
                symmetric = TRUE, only.values = TRUE)$values
# A scree for the compromise
Scree.S <-  PlotScree(ev = ev4S,
                   p.ev = NULL, max.ev = NULL, alpha = 0.05,
                   col.ns = "#006D2C", col.sig = "#54278F",plotKaiser = TRUE,
                   title = "S-mat: Explained Variance per Dimension")

```

### I Set

```{r echo=FALSE}


Fi <- testDistatis$res4Splus$F
col <- brewer.pal(n = 3, name='Set2')

col4Composer <- vector('list',length = nrow(Fi))
col4Composer[grep('Moza', rownames(Fi))] = col[1]
col4Composer[grep('Beet', rownames(Fi))] = col[2]
col4Composer[grep('Bach', rownames(Fi))] = col[3]


#col4Composer <- prettyGraphsColorSelection(nrow(Fi))
#  Use colors from prettyGraphs
#---------------------------------------------------------------------
# Graphs for the I set
#---------------------------------------------------------------------
# Create the base map
constraints4Fi <- lapply(minmaxHelper(Fi),'*',1.2)
baseMap.i <- PTCA4CATA::createFactorMap(Fi,
                                        col.points = col4Composer,
                                        col.labels = col4Composer,
                        constraints = constraints4Fi,
                                        alpha.points =  .4)
#---------------------------------------------------------------------
# We are  interested about the labels here
# so we will use dots and labels
#---------------------------------------------------------------------
# Plain map with color for the I-set
aggMap.i <- baseMap.i$zeMap_background + baseMap.i$zeMap_dots +
                                         baseMap.i$zeMap_text
#---------------------------------------------------------------------
# print this Map
print(aggMap.i)

```


```{r echo=FALSE}

constraints4Fi <- lapply(minmaxHelper(Fi),'*',1.2)
GraphElli <- MakeCIEllipses(BootF[,1:2,],
                      names.of.factors = c("Factor 1","Factor 2"),
                      alpha.line = .5,
                      alpha.ellipse = .3,
                      line.size = .5,
                      line.type = 3,
                      col = unlist(col4Composer),
                      p.level = .95 )
#---------------------------------------------------------------------
# create the I-map with Observations and their confidence intervals
#
aggMap.i.withCI <-  aggMap.i +  GraphElli + MapGroup$zeMap_text
#---------------------------------------------------------------------
# plot it!
print(aggMap.i.withCI)


```