---
title: "All Countries Environmental Data"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---


```{r plsc_init, include=FALSE, results='hide'}
# Installing required packages
# sudo apt-get install libcairo2-dev
# sudo apt-get install libxml2-dev

devtools::install_github('HerveAbdi/PTCA4CATA')
library(corrplot)
library(ExPosition)
library(InPosition)
library(animation)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(gplots)
library(PTCA4CATA)
library(reshape2)

# Source the function file:
file4PLSCfunctions <- '../rmarkdowns/pls/inferences4PLSC.R'
source(file4PLSCfunctions)
#
install.packages('TExPosition')
library(TExPosition)
rm(list = ls())


```

# Partial Least Squares - Correlation

## Description

PLS is used to find the fundamental relations between two matrices (X and Y), i.e. a latent variable approach to modeling the covariance structures in these two spaces. A PLS model will try to find the multidimensional direction in the X space that explains the maximum multidimensional variance direction in the Y space. PLS regression is particularly suited when the matrix of predictors has more variables than observations, and when there is multicollinearity among X values. PLS bears some relation to principal components regression; instead of finding hyperplanes of maximum variance between the response and independent variables, it finds a linear regression model by projecting the predicted variables and the observable variables to a new space. Because both the X and Y data are projected to new spaces, the PLS family of methods are known as bilinear factor models.



```{r plsc_data, include=FALSE, results='hide'}
country_env_df = read.csv('../dataset/World_countries_env_vars.csv')
country_env_df_for_corr = country_env_df[, !names(country_env_df) %in% c('Country')]

country_env_df$Happiness_Rank = cut(country_env_df$Happiness_Rank,breaks = c(0, 39,118, 159), right=TRUE, labels = c('VH' , 'H', 'U'),ordered_result = TRUE)

str(country_env_df)
```



```{r echo=FALSE}

country_env_df_for_pca = country_env_df[, !names(country_env_df) %in% c('Country', 'Happiness_Rank')]

rownames(country_env_df_for_corr) <- paste0(country_env_df[,'Country'], '(' ,country_env_df[,'Happiness_Rank'], ')')
rownames(country_env_df_for_pca) <- paste0(country_env_df[,'Country'], '(' ,country_env_df[,'Happiness_Rank'], ')')

```


## Correlation Plot
Visually analyze multicollinearity between all varaibles in Rain and Temperature tables.  

```{r echo=FALSE}
rain = country_env_df_for_pca[,grep('rain',colnames(country_env_df_for_pca))]
temp = country_env_df_for_pca[,grep('temp',colnames(country_env_df_for_pca))]

corr_result = cor(rain, temp)
corrplot(corr_result,order = 'hclust', addrect = 3)
```


## PLS-C 

```{r plsc, echo=FALSE}

resPLSC <- tepPLS(rain,temp,DESIGN = country_env_df$Happiness_Rank,graphs = FALSE)
resPerm4PLSC <- perm4PLSC(rain, # First Data matrix 
                          temp, # Second Data matrix
                          nIter = 1000 # How mny iterations
                          )
# to see what results we have
print(resPerm4PLSC)


```

Now we have Latent Variables and Saliences. 
* Latent Variables are the new Data points w.r.t. correlation between both the tables. Latent Variables exists for each table.
* Saliences represent correlation between variables of each table. 

## Scree Plot
Gives amount of information explained by corresponding component. Gives an intuition to decide which components best represent data in order to answer the research question.

P.S. The most contribution component may not always be most useful for a given research question. 


```{r echo=FALSE}


PTCA4CATA::PlotScree(ev = resPLSC$TExPosition.Data$eigs,
                      title = 'SCREE Plot',
                      p.ev = resPerm4PLSC$pEigenvalues,
                      plotKaiser = TRUE
)

```

## Latent Variables 

Lets visualize happiness categories for Components 1 of each table

### Component 1 for both Tables: Rain and Temperature
```{r echo=FALSE}

comp1 = data.frame(lx1 = resPLSC$TExPosition.Data$lx[,1], ly1 = resPLSC$TExPosition.Data$ly[,1])
mean_comp1 = comp1
mean_comp1$Happiness_Rank = country_env_df$Happiness_Rank

list_color <- c('U' = 'burlywood3', 'H' = 'royalblue1', 'VH' = 'springgreen4')

lx_col <- recode (country_env_df$Happiness_Rank, !!!list_color)

country_env_mean_df = aggregate(comp1,by = list(country_env_df$Happiness_Rank), mean)
rownames(country_env_mean_df) <- c('Very Happy', 'Happy', 'Unhappy')

country_env_mean_df <- country_env_mean_df[,2:3]

country_factor_map <- PTCA4CATA::createFactorMap(comp1, title='', 
                                                 col.points = lx_col,
                                                 #col.labels = country_env_df$Happiness_Rank,
                                                 
                                                 display.labels = FALSE)

country_factor_mean_map <- PTCA4CATA::createFactorMap(country_env_mean_df, title='', 
                                                 col.points = c('springgreen4', 'royalblue1', 'burlywood3'),
                                                 col.labels = c('springgreen4', 'royalblue1', 'burlywood3'),
                                                 cex = 8,alpha.points = 0.8,
                                                 display.labels = TRUE)

country_label4Map <- PTCA4CATA::createxyLabels.gen(1,2,lambda = resPLSC$TExPosition.Data$eigs,  tau = resPLSC$TExPosition.Data$t,axisName = "Table") 


country_map = country_factor_map$zeMap + country_label4Map + country_factor_mean_map$zeMap_dots + country_factor_mean_map$zeMap_text + ggtitle('Component 1')

print(country_map)


```

### Component 2 for both Tables: Rain and Temperature
```{r echo=FALSE}

comp1 = data.frame(lx1 = resPLSC$TExPosition.Data$lx[,2], ly1 = resPLSC$TExPosition.Data$ly[,2])
mean_comp1 = comp1
mean_comp1$Happiness_Rank = country_env_df$Happiness_Rank

list_color <- c('U' = 'burlywood3', 'H' = 'royalblue1', 'VH' = 'springgreen4')

lx_col <- recode (country_env_df$Happiness_Rank, !!!list_color)

country_env_mean_df = aggregate(comp1,by = list(country_env_df$Happiness_Rank), mean)
rownames(country_env_mean_df) <- c('Very Happy', 'Happy', 'Unhappy')

country_env_mean_df <- country_env_mean_df[,2:3]

 country_factor_map <- PTCA4CATA::createFactorMap(comp1, title='', 
                                                 col.points = lx_col,
                                                 #col.labels = country_env_df$Happiness_Rank,
                                                 
                                                 display.labels = FALSE)

country_factor_mean_map <- PTCA4CATA::createFactorMap(country_env_mean_df, title='', 
                                                 col.points = c('springgreen4', 'royalblue1', 'burlywood3'),
                                                 col.labels = c('springgreen4', 'royalblue1', 'burlywood3'),
                                                 cex = 8,alpha.points = 0.8,
                                                 display.labels = TRUE)

country_label4Map <- PTCA4CATA::createxyLabels.gen(1,2,lambda = resPLSC$TExPosition.Data$eigs, tau = resPLSC$TExPosition.Data$t, axisName = 'Table') 


country_map = country_factor_map$zeMap + country_label4Map + country_factor_mean_map$zeMap_dots + country_factor_mean_map$zeMap_text + ggtitle('Component 2')

print(country_map)


```



## Salience for Rain

### Components 1

```{r echo=FALSE}

PrettyBarPlot2(resPLSC$TExPosition.Data$fi[,1],
              threshold = 0,
              font.size = 5,
              main = 'Salience for Rain',
              horizontal = TRUE,
              angle.text = 90,
              ylim=c(-3,3)
              )

```

### Component 2

```{r echo=FALSE}

PrettyBarPlot2(resPLSC$TExPosition.Data$fi[,2],
              threshold = 0,
              font.size = 5,
              main = 'Salience for Rain',
              horizontal = TRUE,
              angle.text = 90,
              ylim=c(-3,3)
              )

```


## Salience for Temperature


### Component 1
```{r echo=FALSE}

PrettyBarPlot2(resPLSC$TExPosition.Data$fj[,1],
              threshold = 0,
              font.size = 4,
              main = 'Salience for Temperature, Component 1',
              horizontal = TRUE,
              angle.text = 90
              )

```

### Component 2
```{r echo=FALSE}

PrettyBarPlot2(resPLSC$TExPosition.Data$fj[,2],
              threshold = 0,
              font.size = 4,
              main = 'Salience for Temperature',
              horizontal = TRUE,
              angle.text = 90
              )

```


## Most Contributing Variables - PLS-C (with Inference)


```{r echo=FALSE }

resBoot4PLSC <- Boot4PLSC(rain, # First Data matrix 
                          temp, # Second Data matrix
                          nIter = 1000, # How many iterations
                      Fi = resPLSC$TExPosition.Data$fi,
                      Fj = resPLSC$TExPosition.Data$fj,
                      nf2keep = 3,
                      critical.value = 2,
                      # To be implemented later
                      # has no effect currently
                      alphaLevel = .05)
#
# to see what results we have
print(resBoot4PLSC)

```


### Bootstrap Test

* Rain - Component 1

```{r echo=FALSE}

BR = resBoot4PLSC$bootRatios.i

PrettyBarPlot2(BR[,1],
              threshold = 2,
              font.size = 5,
              main = 'Bootstrap ratio For Rain ',
              ylab = 'Bootstrap ratios',
              horizontal = TRUE,
              ylim = c(-10,12)
              )

```


* Rain - Component 2

```{r echo=FALSE}

BR = resBoot4PLSC$bootRatios.i

PrettyBarPlot2(BR[,2],
              threshold = 2,
              font.size = 5,
              #color4bar = gplots::col2hex(col4J), # we need hex code
              main = 'Bootstrap ratio For Rain ',
              ylab = 'Bootstrap ratios',
              horizontal = TRUE,
              ylim = c(-10,12)
              )

```


* Temperature - Component 1

```{r echo=FALSE}

BR = resBoot4PLSC$bootRatios.j

PrettyBarPlot2(BR[,1],
              threshold = 2,
              font.size = 4,
              #color4bar = gplots::col2hex(col4J), # we need hex code
              main = 'Bootstrap ratio For Temperature ',
              ylab = 'Bootstrap ratios',
              horizontal = TRUE,
              ylim = c(-20,15)
              )

```




* Temperature - Component 2

```{r echo=FALSE}

BR = resBoot4PLSC$bootRatios.j

PrettyBarPlot2(BR[,2],
              threshold = 2,
              font.size = 4,
              #color4bar = gplots::col2hex(col4J), # we need hex code
              main = 'Bootstrap ratio For Temperature ',
              ylab = 'Bootstrap ratios',
              horizontal = TRUE,
              ylim = c(-20,15)
              )

```

## Conclusion

Here Component 2 seems to best seperate Happiness levels. Let's compare Component 2 for both tables.

* Table 1 & 2 Component 2
  + Latent Variables: Very Happy vs Unhappy (for Rain and Temperature both)
  + Salience:
    * Rain: It seems dryness and wetness at a montly scale have more effect than coldness or yearly patterns.
    * Temperature: All temperature variations at a monthly and yearly scale seems to impact happiness.
  

